---
title: "R Notebook"
output: html_notebook
---

# Load tidyverse 

```{r}
library(corrr)
```


```{r}
library(tidyverse)
```
```{r}
library(readxl)
```

```{r}
library(tidyjson)
```
```{r}
library(feather)
```

```{r}
library(ggExtra)
```

```{r}
library(cowplot)
```

```{r}
library(arrow)
```

```{r}
library(MetBrewer)
```


# Load data

```{r}
df <- arrow::read_feather("../data/features_combined_20220201.feather")
```

```{r}
df %>% head()
```


# Explore data

### Check if scores and tfm classification NaNs align

```{r}
df %>% 
  mutate(empty_border = is.na(Border_Scores),
         empty_asymmetry = is.na(Asymmetry_Scores),
         empty_tfm = is.na(TFM_predictions),
         empty_match = empty_tfm == (empty_border | empty_asymmetry)) %>% 
  count(empty_match)
```
```{r}
df %>% 
  mutate(empty_border = is.na(Border_Scores),
         empty_asymmetry = is.na(Asymmetry_Scores),
         empty_tfm = is.na(TFM_predictions),
         empty_match = empty_tfm == (empty_border | empty_asymmetry)) %>% 
  filter(!empty_match)
```
```{r}
df %>% 
  mutate(empty_tfm = is.na(TFM_predictions)) %>% 
  count(empty_tfm)
```
```{r}
df %>% 
  mutate(empty_tfm = is.na(TFM_probability_melanoma)) %>% 
  count(empty_tfm)
```

```{r}
df %>% 
  mutate(empty_border = is.na(Border_Scores),
         empty_asymmetry = is.na(Asymmetry_Scores),
         empty_tfm = is.na(TFM_probability_melanoma),
         empty_match = empty_tfm == (empty_border | empty_asymmetry)) %>% 
  count(empty_match)
```
```{r}
df %>% colnames()
```

# Analyse data

```{r}
df %>% 
  count(observation_type)
```


## Create a predictive performance variable that captures the difference between model predictions and true classifications (square error of the probabilities)
```{r}
null_prediction <- df %>% 
  filter(observation_type == "train") %>% 
  pull(ground_truth) %>% 
  mean(na.rm = T)
```


```{r}
df <- df %>% 
  mutate(E2_TFM = (TFM_probability_melanoma - ground_truth)**2,
         E2_DNN = (DNN_probability_melanoma - ground_truth)**2,
         E2_baseline = (0.5 - ground_truth)**2)
```

```{r}
df %>% 
  filter(observation_type == "test") %>% 
  mutate(acc_TFM = round(TFM_probability_melanoma) == ground_truth,
         acc_DNN = round(DNN_probability_melanoma) == ground_truth) %>% 
  summarise(mean(acc_TFM, na.rm = T), mean(acc_DNN))
```


```{r}
df <- df %>% 
  filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores) &
        border_ratio > 0.5)
```
```{r}
library(ggbeeswarm)
```

```{r}
baseline_value <- df %>% 
  filter(observation_type == "test") %>% 
  summarise(mean(E2_baseline)) %>% 
  pull()
baseline_value
```
```{r}
 df %>% 
  filter(observation_type == "test") %>% 
  summarise(mean(E2_DNN)) %>% 
  pull()
```
```{r}
 df %>% 
  filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  summarise(mean(E2_TFM, na.rm = T)) %>% 
  pull()
```

```{r}
squared_error_comparison <- df %>%  
  filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  select(image_id, E2_TFM, E2_DNN) %>% 
  pivot_longer(cols = -image_id, names_to = "Model", values_to = "E2") %>%
  #group_by(Model) %>% 
  #summarise(mean(E2))
  ggplot(aes(y = E2, x = Model)) +
    stat_summary(fun.data = "mean_cl_normal", color = "black") +
    geom_hline(yintercept=baseline_value, , size =1.2, color = "gray40", linetype ="longdash") +
  coord_cartesian(ylim = c(0, 0.26))  +
  ylab("Squared error") +
  scale_x_discrete(labels = c("DNN", "TFM")) +
  theme_minimal()
squared_error_comparison
```


### Figure 2 two versions

```{r}
e2_scatter <- df %>% 
    filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = E2_TFM, y = E2_DNN, color = factor(ground_truth), shape = factor(ground_truth))) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma absent", "Melanoma present"), name = "") +
  scale_shape_manual(values = c(19, 15), labels =c("Melanoma absent", "Melanoma present"), name = "") +
  xlim(0, 1) +
  ylim(0, 1) +
  labs(x = "TFM Squared error", y = "DNN squared error") +
  theme_minimal() +
  theme(legend.position = c(0.8, 1.15))
e2_scatter
```

```{r}
hist_dist_e2_tfm <- df %>% 
        filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = E2_TFM, fill = factor(ground_truth))) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  scale_shape_manual(values = c(19, 15), labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  geom_histogram(alpha = 0.5, bins = 20) +
  scale_fill_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "", guide = "none") +
  xlim(0, 1) +
  ylim(0, 55) +
  labs(x = "", y = "") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_blank(),
        axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
hist_dist_e2_tfm
```

```{r}
hist_dist_e2_dnn <- df %>% 
          filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = E2_DNN, fill = factor(ground_truth))) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  geom_histogram(alpha = 0.5, bins = 20) +
  scale_fill_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "", guide = "none") +
  xlim(0, 1) +
  ylim(0, 55) +
  labs(x = "", y = "") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_blank(),
        axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank()) +
  coord_flip()
```

```{r}
hist_dist_e2_dnn
```

```{r}
combined_plot_1 <- insert_xaxis_grob(e2_scatter, hist_dist_e2_tfm, position = "top")
combined_plot_2 <- insert_yaxis_grob(combined_plot_1, hist_dist_e2_dnn, position = "right")
ggsave(plot = combined_plot_2, filename = "../output/Fig2_Squared_Error.pdf", width = 10, height = 10, units = "cm")
```
```{r}
prob_scatter <- df %>% 
    filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = TFM_probability_melanoma, y = DNN_probability_melanoma, color = factor(ground_truth), shape = factor(ground_truth))) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Benign", "Melanoma"), name = "") +
  scale_shape_manual(values = c(19, 15), labels =c("Benign", "Melanoma"), name = "") +
  xlim(0, 1) +
  ylim(0, 1) +
  labs(x = "TFM probability malign", y = "DNN probability malign") +
  theme_minimal() +
  theme(legend.position = c(0.97, 1.18))
prob_scatter
```

```{r}
hist_dist_prob_tfm <- df %>% 
        filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = TFM_probability_melanoma, fill = factor(ground_truth))) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  scale_shape_manual(values = c(19, 15), labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  geom_histogram(alpha = 0.5, bins = 20) +
  scale_fill_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "", guide = "none") +
  xlim(0, 1) +
  ylim(0, 55) +
  labs(x = "", y = "") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_blank(),
        axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
hist_dist_prob_tfm
```

```{r}
hist_dist_prob_dnn <- df %>% 
          filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  ggplot(aes(x = DNN_probability_melanoma, fill = factor(ground_truth))) +
  scale_color_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "") +
  geom_histogram(alpha = 0.5, bins = 20) +
  scale_fill_manual(values = met.brewer("Troy", 8)[c(7, 4)], labels =c("Melanoma benign", "Melanoma malign"), name = "", guide = "none") +
  xlim(0, 1) +
  ylim(0, 55) +
  labs(x = "", y = "") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_blank(),
        axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank()) +
  coord_flip()
```

```{r}
hist_dist_prob_dnn
```

```{r}
combined_plot_1 <- insert_xaxis_grob(prob_scatter, hist_dist_prob_tfm, position = "top")
combined_plot_2 <- insert_yaxis_grob(combined_plot_1, hist_dist_prob_dnn, position = "right")
```
# Save figure 2

```{r}
fig2 <- plot_grid(squared_error_comparison, combined_plot_2, labels = "AUTO", rel_widths = c(1, 2))
```
```{r}
ggsave("../output/Fig2_tfm_vs_dnn.pdf", plot = fig2, width = 13, height = 8, units = "cm")
```

# Statistical analysis

## Log likelihood ratios

```{r}
bernoulli_likelihood = function(f,x){ f^x*(1-f)^(1-x) }
```

```{r}
(likelihood_table <- df %>%  
  filter(observation_type == "test" &
        !is.na(Asymmetry_Scores) &
        !is.na(Border_Scores)) %>% 
  mutate(likelihood_null = bernoulli_likelihood(0.5, ground_truth),
         likelihood_TFM = bernoulli_likelihood(TFM_probability_melanoma, ground_truth),
         likelihod_DNN = bernoulli_likelihood(DNN_probability_melanoma, ground_truth)) %>% 
  summarise(null = prod(likelihood_null),
            TFM = prod(likelihood_TFM),
            DNN = prod(likelihod_DNN)) %>% 
  pivot_longer(c(null, TFM, DNN), names_to = "model", values_to = "likelihood") %>% 
  mutate(log_likelihood = log(likelihood)))
```
```{r}
null_log_likelihood <- likelihood_table %>% filter(model == "null") %>% pull(log_likelihood)
tfm_log_likelihood <- likelihood_table %>% filter(model == "TFM") %>% pull(log_likelihood)
dnn_log_likelihood <- likelihood_table %>% filter(model == "DNN") %>% pull(log_likelihood)
```


```{r}
round(tfm_log_likelihood - null_log_likelihood, 2)	
```

### See how to compute a p-value from a log-likelihood ratio using the chi-squared distribution: https://www.statisticshowto.com/likelihood-ratio-tests/



```{r}
pchisq(-2*(null_log_likelihood - tfm_log_likelihood), df = 4, lower.tail = FALSE)
```
```{r}
-2*(tfm_log_likelihood - null_log_likelihood)
```


```{r}
round(dnn_log_likelihood - tfm_log_likelihood, 2)	
```
```{r}
round(dnn_log_likelihood - null_log_likelihood, 2)	
```


```{r}
cor.test(~ TFM_probability_melanoma + DNN_probability_melanoma, data = df)
```

